### 2019-03-14 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Done for now
Added experiments about choice of threshold functions, and choice of exploration mechanism.
Sorry Nabil I couldn't be there...
Almost done for this chapter 6. Maybe I'll move all these « exploration » numerical experiments in the main text, but use smaller tables (explore less values?)

%-----------------------------------------------------------------------------
\paragraph{False alarm probability.}

@@

Lemma~\ref{lem:6:FalseAlarm} provides the first control of false alarm for the B-GLRT employed for bounded data. However, the threshold \eqref{def:6:beta} is not fully explicit as the function $\cT(x)$ can only be computed numerically.
Note that for sub-Gaussian distributions, results from \cite{Maillard2018GLR} show that the smaller and more explicit threshold
$\beta(n,\delta) = \left(1 + \frac{1}{n}\right)\ln\left(\frac{3n\sqrt{n}}{\delta}\right)$,
can be used to prove an upper bound of $\delta$ for the false alarm probability of the GLR, with quadratic divergence $\kl(x,y)=2(x-y)^2$.
%
For the B-GLRT, numerical simulations suggest that the threshold \eqref{def:6:beta} is a bit conservative (see Appendix~\ref{sec:6:exploringDifferentThresholdFunctions}), and in practice we recommend to keep only the leading term and use $\beta(n,\delta) = \ln(3n\sqrt{n}/\delta)$.

@@

The \GLRklUCB{} algorithm can be viewed as a \klUCB{} algorithm allowing for some \emph{restarts} on the different arms. A restart happens when the B-GLRT change-point detector detects a change on the arm that has been played (line $9$).
%
To be fully specific, $\GLR_\delta(Z_1,\dots,Z_n) = \mathrm{True}$ if and only if
\begin{equation}
    \sup_{1 < s < n} \left[s \times \kl \left(\frac{1}{s}\sum_{i=1}^s Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right) + (n-s) \times \kl \left(\frac{1}{n-s}\sum_{i=s+1}^t Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right)\right] \geq \beta(n,\delta),
\end{equation}
%
with $\beta(n,\delta)$ defined in \eqref{def:6:beta}, or $\beta(n,\delta) = \ln(3n^{3/2}/\delta)$, as recommended in practice, see Section~\ref{sub:6:PracticalConsiderations}. We define the (\klUCB{} like) index used by our algorithm, by denoting
$\tau_i(t)$ the last restart that happened for arm $i$ before time $t$,
\[n_i(t) = \sum_{s=\tau_i(t)+1}^{t} \indic(A_s=i)\]
the number of selections of arm $i$, and
\[\widehat{\mu}_i(t) = \frac{1}{n_i(t)} \sum_{s=\tau_i(t)+1}^{t}X_{i,s} \indic(A_s=i)\]
their empirical mean (if $n_i(t)\neq0$).
%
With the exploration function $f(t) = \ln(t) + 3 \ln(\ln(t))$ if $t>1$ and $f(t)=0$ otherwise,
the index is defined using the binary relative entropy as
%
\begin{equation}\label{def:6:UCB}
    \UCB_i(t) := \max \Bigl\{ q\in[0,1] : n_i(t) \times \kl\left(\widehat{\mu}_i(t),q\right) \leq f(t - \tau_i(t)) \Bigr\}.
\end{equation}

@@

We present in Appendix~\ref{sec:6:exploringDifferentThresholdFunctions} numerical simulations that compare different choices of threshold functions $\beta$.
We compare the complicated value used in Lemma~\ref{lem:6:FalseAlarm}, using a numerical approximation of the function $\cT$,
with the simpler value $\beta(n,\delta) = \ln(3n\sqrt{n}/\delta)$,
as well as a few other choices.
The sum-up of these simulations is that it is enough to use a simpler and more explicit threshold to have an efficient change-point detection algorithm, and an efficient \GLRklUCB{} policy.

@@

Recall that the regret is defined as $R_T = \bE\left[\sum_{t=1}^T(\mu_{i^*_t}(t) - \mu_{I_t}(t))\right]$.
Introducing $\cD(T,\alpha)$ the (deterministic) set of time steps at which the forced exploration is performed before time $T$ (see lines $3$-$4$ in Algorithm~\ref{algo:6:GLRklUCB}), one can write, using notably that $\mu_{i^*_t} - \mu_{I_t} \leq 1$, due to the bounded rewards,
\begin{align*}
 & \sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t)) \\
 & \hspace{1cm} \leq  \sum_{t=1}^{T} \indic(t \in \cD(T,\alpha)) + \sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(t \notin\cD(T,\alpha), \UCB_{I_t}(t) \geq \UCB_{i^*_t}(t)\right) \\
 & \hspace{1cm}\leq  \alpha T + \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right) + \sum_{i=1}^K\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{i}(t))\indic\left(I_t = i, \UCB_{i}(t) \geq \mu_{i^*_t}(t)\right).
\end{align*}
%
Introducing some \emph{good event} $\cE_T$, to be specified in each case, and its complementary $\cE_T^c$, one can write the following decomposition, that highlights two important terms $(A)$ and $(B)$
%
\begin{align}\label{eq:6:GeneRegretBound}
    R_T \leq T \bP\left(\cE_T^c\right) + \alpha T & + \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} \\
    & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}. \nonumer
\end{align}
%
Each analysis requires to define an \emph{appropriate good event}, stating that \emph{some} change-points are detected within a reasonable delay. Each regret bound then follows from upper bounds on term $(A)$, term $(B)$, and on the failure probability $\bP(\cE_T^c)$.
%
To control $(A)$ and $(B)$, we split the sum over consecutive segments $[\tau^{(k)}+1, \tau^{(k+1)}]$ for global changes and $[\tau^{(k)}_i+1, \tau^{(k+1)}_i]$ for each arm $i$ for local changes, and use elements from the analysis of \klUCB{} of \cite{KLUCBJournal}.

The tricky part of each proof, which crucially exploits Assumption~\ref{ass:6:LongPeriodsGlobal} or \ref{ass:6:LongPeriods}, is actually to obtain an upper bound on $\bP(\cE_T^c)$. For example for local changes (Theorem~\ref{thm:6:mainRegretBound}), the good event is defined as
\begin{equation}
    \cE_T(\alpha,\delta) = \left(\forall i \in \{1, \ldots, K\}, \forall \ell \in \{1, \ldots, \NCi\}, \hat{\tau}^{(\ell)}_i \in \left[\tau_i^{(\ell)} + 1, \tau_i^{(\ell)} + d_i^{(\ell)}\right] \right),
\end{equation}
%
where $\widehat{\tau}_i^{(\ell)}$ is defined as the $\ell$-th change detected by the algorithm on arm $i$ and $d_i^{(\ell)}=d_i^{(\ell)}(\alpha,\delta)$ is defined in Assumption~\ref{ass:6:LongPeriods}.
Introducing the event
\[\cC_i^{(\ell)} = \left\{\forall j \leq \ell, \widehat{\tau}_i^{(\ell)} \in \left[\tau_i^{(j)} + 1, \tau_i^{(j)} + d_i^{(j)} \right] \right\},\]
that all the changes up to the $\ell$-th have been detected, a union bound yields the following decomposition
%
\begin{align}
    \bP(\cE_T(\alpha,\delta)^c) & \leq \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\widehat{\tau}_i^{(\ell)} \leq \tau_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(a)} + \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\widehat{\tau}_i^{(\ell)} \geq \tau_i^{(\ell)} + d_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(b)}.
\end{align}

@@

For \GLRklUCB, we explore the two different options with \textbf{Local} and \textbf{Global} restarts,
using respectively
$\delta = 1/\sqrt{\Upsilon T}$, $\alpha = \alpha_0 \sqrt{\Upsilon_T \log(T) /T}$
and
$\delta = 1/\sqrt{K \Upsilon_T T}$, $\alpha = \alpha_0\sqrt{K \Upsilon_T \log(T) / T}$
from Corollaries \ref{cor:6:Global} and \ref{cor:6:Local}.
Both choices of $\alpha$ appear to be too large empirically, as they come from minimizing the regret upper-bound rather than the regret itself, thus
the constant is set to $\alpha_0 = 0.05$
(we show in Appendix~\ref{sec:6:choosingAlpha0} a certain robustness, with similar regret as soon as $\alpha\leq0.1$).
%
To speed up our simulations, two optimizations are used, with $\Delta n = \Delta s = 10$,
and \CUSUM{} also uses the first trick with $\Delta n = 10$ (see Appendix~\ref{sub:6:IdeasOptimizations} for more details).

>>>>

%-----------------------------------------------------------------------------
\paragraph{False alarm probability.}\label{par:6:falseAlarm}

@@

Lemma~\ref{lem:6:FalseAlarm} provides the first control of false alarm for the B-GLRT employed for bounded data. However, the threshold \eqref{def:6:beta} is not fully explicit as the function $\cT(x)$ can only be computed numerically.
Note that for sub-Gaussian distributions, results from \cite{Maillard2018GLR} show that the smaller and more explicit threshold
$\beta(n,\delta) = \left(1 + \frac{1}{n}\right)\ln\left(\frac{3n\sqrt{n}}{\delta}\right)$,
can be used to prove an upper bound of $\delta$ for the false alarm probability of the GLR, with quadratic divergence $\kl(x,y)=2(x-y)^2$.
%
For the B-GLRT, numerical simulations suggest that the threshold \eqref{def:6:beta} is a bit conservative (see Appendix~\ref{sec:6:exploringDifferentThresholdFunctions}), and in practice we recommend to keep only the leading term and use $\beta(n,\delta) = \ln(3n^{3/2}/\delta) = \ln(3) + 3/2\log(n) - \log(\delta)$.

@@

The \GLRklUCB{} algorithm can be viewed as a \klUCB{} algorithm allowing for some \emph{restarts} on the different arms. A restart happens when the B-GLRT change-point detector detects a change on the arm that has been played (line $9$).
%
To be fully specific, $\GLR_\delta(Z_1,\dots,Z_n) = \mathrm{True}$ if and only if
\begin{equation}\label{eq:whenDoesGLRklUCBrestart}
    \sup_{1 < s < n} \left[s \times \kl \left(\frac{1}{s}\sum_{i=1}^s Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right) + (n-s) \times \kl \left(\frac{1}{n-s}\sum_{i=s+1}^t Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right)\right] \geq \beta(n,\delta),
\end{equation}
%
with $\beta(n,\delta)$ defined in \eqref{def:6:beta}, or $\beta(n,\delta) = \ln(3n^{3/2}/\delta)$, as recommended in practice, see Section~\ref{sub:6:PracticalConsiderations}. We define the (\klUCB{} like) index used by our algorithm, by denoting
$\tau_i(t)$ the last restart that happened for arm $i$ before time $t$,
$n_i(t) = \sum_{s=\tau_i(t)+1}^{t} \indic(A_s=i)$
the number of selections of arm $i$, and
$\widehat{\mu}_i(t) = \frac{1}{n_i(t)} \sum_{s=\tau_i(t)+1}^{t}X_{i,s} \indic(A_s=i)$
their empirical mean (if $n_i(t)\neq0$).
%
With the exploration function $f(t) = \ln(t) + 3 \ln(\ln(t))$ if $t>1$ and $f(t)=0$ otherwise,
the index is defined using the binary relative entropy as
%
\begin{equation}\label{def:6:UCB}
    \UCB_i(t) := \max \Bigl\{ q\in[0,1] : n_i(t) \times \kl\left(\widehat{\mu}_i(t),q\right) \leq f(t - \tau_i(t)) \Bigr\}.
\end{equation}

@@

We present in Appendix~\ref{sec:6:exploringDifferentThresholdFunctions} numerical simulations that compare different choices of threshold functions $\beta$.
We compare the complicated value used in Lemma~\ref{lem:6:FalseAlarm}, using a numerical approximation of the function $\cT$,
with the simpler value $\beta(n,\delta) = \ln(3n\sqrt{n}/\delta)$, as well as two other choices.
The sum-up of these simulations is that it is enough to use a simpler and more explicit threshold, thus we recommend in practice to use $\beta_1(n,\delta) = \ln(3n\sqrt{n}/\delta)$.
%  to have an efficient change-point detection algorithm, and an efficient \GLRklUCB{} policy.

@@

Recall that the regret is defined as $R_T = \bE\left[\sum_{t=1}^T(\mu_{i^*_t}(t) - \mu_{I_t}(t))\right]$.
Introducing $\cD(T,\alpha)$ the (deterministic) set of time steps at which the forced exploration is performed before time $T$ (see lines $3$-$4$ in Algorithm~\ref{algo:6:GLRklUCB}),
we observe that we have $\mu_{i^*_t} - \mu_{I_t} \leq 1$ because rewards are assumed to be bounded in $[0,1]$,
and thus we can write a first decomposition of the regret,
\begin{align*}
 & \sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t)) \\
 & \hspace{1cm} \leq  \sum_{t=1}^{T} \indic(t \in \cD(T,\alpha)) + \sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(t \notin\cD(T,\alpha), \UCB_{I_t}(t) \geq \UCB_{i^*_t}(t)\right) \\
 & \hspace{1cm}\leq  \alpha T + \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right) + \sum_{i=1}^K\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{i}(t))\indic\left(I_t = i, \UCB_{i}(t) \geq \mu_{i^*_t}(t)\right).
\end{align*}
%
Introducing some \emph{good event} $\cE_T$, to be specified in each case, and its complementary $\cE_T^c$, one can write the following decomposition, that highlights two important terms $(A)$ and $(B)$
%
\begin{align}\label{eq:6:GeneRegretBound}
    R_T \leq T \bP\left(\cE_T^c\right) + \alpha T & + \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} \\
    & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}. \nonumber
\end{align}
%
Each analysis requires to define an \emph{appropriate good event}, stating that \emph{some} change-points are detected within a reasonable delay. Each regret bound then follows from upper bounds on term $(A)$, term $(B)$, and on the failure probability $\bP(\cE_T^c)$.
%
To control $(A)$ and $(B)$, we split the sum over consecutive segments $[\tau^{(k)}+1, \tau^{(k+1)}]$ for global changes and $[\tau^{(k)}_i+1, \tau^{(k+1)}_i]$ for each arm $i$ for local changes, and use elements from the analysis of \klUCB{} of \cite{KLUCBJournal}.

The tricky part of each proof, which crucially exploits Assumption~\ref{ass:6:LongPeriodsGlobal} or \ref{ass:6:LongPeriods}, is actually to obtain an upper bound on $\bP(\cE_T^c)$. For example for local changes (Theorem~\ref{thm:6:mainRegretBound}), the good event is defined as
\begin{equation}
    \cE_T(\alpha,\delta) = \left(\forall i \in \{1, \ldots, K\}, \forall \ell \in \{1, \ldots, \NCi\}, \hat{\tau}^{(\ell)}_i \in \left[\tau_i^{(\ell)} + 1, \tau_i^{(\ell)} + d_i^{(\ell)}\right] \right),
\end{equation}
%
where $\widehat{\tau}_i^{(\ell)}$ is defined as the $\ell$-th change detected by the algorithm on arm $i$ and $d_i^{(\ell)}=d_i^{(\ell)}(\alpha,\delta)$ is defined in Assumption~\ref{ass:6:LongPeriods}.
Introducing the event
\[\cC_i^{(\ell)} = \left\{\forall j \leq \ell, \widehat{\tau}_i^{(\ell)} \in \left[\tau_i^{(j)} + 1, \tau_i^{(j)} + d_i^{(j)} \right] \right\},\]
that all the changes up to the $\ell$-th have been detected, a union bound yields this decomposition
%
\begin{align}
    \bP(\cE_T(\alpha,\delta)^c) & \leq \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\widehat{\tau}_i^{(\ell)} \leq \tau_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(a)} + \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\widehat{\tau}_i^{(\ell)} \geq \tau_i^{(\ell)} + d_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(b)}.
\end{align}

@@

For \GLRklUCB, we explore the two different options with \textbf{Local} and \textbf{Global} restarts,
using respectively
$\delta = 1/\sqrt{\Upsilon T}$, $\alpha = \alpha_0 \sqrt{\Upsilon_T \log(T) /T}$
and
$\delta = 1/\sqrt{K \Upsilon_T T}$, $\alpha = \alpha_0\sqrt{K \Upsilon_T \log(T) / T}$
from Corollaries \ref{cor:6:Global} and \ref{cor:6:Local}.
Both choices of $\alpha$ appear to be too large empirically, as they come from minimizing the regret upper-bound rather than the regret itself, thus
the constant is set to $\alpha_0 = 0.05$.
We show in Appendix~\ref{sec:6:choosingAlpha0} a certain robustness, with similar regret as soon as $\alpha\leq0.1$.
We do not use a constant $\delta_0$ to change the confidence level to $\delta = \delta_0 \delta_T$, as we show in Appendix~\ref{sec:6:choosingDelta} that $\delta_0=1$ is uniformly better for different problems.
%
To speed up our simulations, two optimizations are used, with $\Delta n = \Delta s = 10$,
and \CUSUM{} also uses the first trick with $\Delta n = 10$ (see Appendix~\ref{sub:6:IdeasOptimizations} for more details).
